---
output:
  pdf_document: default
  html_document: default
---
#Analysis of Recombination rate variation
###written by Ulku Huma Altindag

##Set directory you are working in with the datasets
```{r}
#setwd(~/PC/Desktop/plasticity experiments/Experiment 4)
```

##load the datasets
```{r}
yv=read.csv("Exp3_rawdata.csv", header= TRUE, stringsAsFactors = TRUE, na.strings="-")
backcross=read.csv("Exp3bc_setup.csv", header=T,na.strings="-")
```

 
##to make sure R reads it as a character rather than a number.

```{r}
backcross$Treatment=as.character(backcross$Treatment)
yv$wildtype=as.numeric(as.character(yv$wildtype))
yv$yellow.vermillion=as.numeric(as.character(yv$yellow.vermillion))
yv$yellow.only=as.numeric(as.character(yv$yellow.only))
yv$vermillion.only=as.numeric(as.character(yv$vermillion.only))
yv$wildtype.1=as.numeric(as.character(yv$wildtype.1))
yv$yellow.vermillion.1=as.numeric(as.character(yv$yellow.vermillion.1))
yv$yellow.only.1=as.numeric(as.character(yv$yellow.only.1))
yv$vermillion.only.1=as.numeric(as.character(yv$vermillion.only.1))
```

##necessary packages
```{r}
library(ggplot2)
library(ggthemes)
library(devtools)
library(gridExtra)
library(gtable)
library(grid)
library(lsmeans)
library(lme4)
library(lmerTest)
library(doBy)
library(pbkrtest)
library(tidyr)
library(reshape2)
library(car)
```
##CO groups were defined

```{r}
yv$SCO1=yv$yellow.only+yv$yellow.only.1
yv$SCO2=yv$vermillion.only+yv$vermillion.only.1
yv$NCO1=yv$wildtype+yv$wildtype.1
yv$NCO2=yv$yellow.vermillion+yv$vermillion.only.1
```
##Using the sanity check, we can see whether R interprets the results accurately.

```{r}
sco_count=sum(yv$SCO1, na.rm = TRUE)+sum(yv$SCO2, na.rm = TRUE)
nco_count=sum(yv$NCO1+yv$NCO2, na.rm = TRUE)
num_samples=sum(nco_count+sco_count, na.rm = TRUE)
```

Rate between yellow and vermillion is ` {r} (sum(yv$SCO1, na.rm = TRUE)+sum(yv$SCO2, na.rm = TRUE))/num_samples`

##merge with treatment data (This is important for summarization of the data)

```{r}
yv_merged <- merge(yv, backcross, by.x = "Vial..", by.y = "Vial.number", all=T)
yv_merged = na.omit(yv_merged)
dataset=summaryBy(SCO1+SCO2+NCO1+NCO2~F1.Vial+Day..letter.of.vial.+Treatment,data=yv_merged, FUN=sum,na.rm=T)
```
##calculating the fecundity
##add in a column for total offspring

```{r}
dataset$total_offspring=dataset$SCO1.sum + dataset$SCO2.sum + dataset$NCO1.sum + dataset$NCO2.sum
```
##add a column for mothers from the F1 generation
```{r}
num_moms=vector(mode="numeric",length=length(dataset$F1.Vial))
```

##loop through dataset to get the count
```{r}
for (h in 1:length(dataset$F1.Vial)) { 
  f1_vial=dataset$F1.Vial[h]
#  f1_vial=as.numeric(gsub("[^0-9\\.]","",f1_vial))
#  print(f1_vial)
  mom_ct=length(unique(sort(subset(yv_merged,yv_merged$F1.Vial==f1_vial)$Vial..)))
  #store result in the vector
  num_moms[h]=mom_ct
}
dataset$Num_moms=num_moms
```
##fecundity equals to total # of offsprings divided by # of mothers

```{r}
dataset$fecundity=dataset$total_offspring/dataset$Num_moms
```

##producing a cleanup version of the datasets

```{r}
write.csv(dataset,file="Experiment3_data.csv")
```

##Fecundity_figure

```{r}
Fecund_figure=ggplot(aes(y=fecundity,x=Day..letter.of.vial., col=Treatment,label=Num_moms),data=dataset)+ylab("# Progeny per mom")+ggtitle("Total Fecundity vs. Days post-mating")+theme_base()

Fecund_figure=Fecund_figure+scale_colour_manual(values=c("blue", "red"))+geom_point(alpha=0.6,size=3)+scale_x_discrete(name="Day",labels=c("1-2","3-4","5-6","7-9","10-12","13-15"))
Fecund_figure=Fecund_figure+stat_summary(fun.y = median, geom="line",aes(group=Treatment),size=2)+geom_text(check_overlap = F,hjust = 0, nudge_x = 0.05,angle=45,size=3)
Fecund_figure
```

##remove vials with few progeny

```{r}
dataset=dataset[dataset$total_offspring>=10,]
```

##Calculate the recombination rate for each replicate

```{r}
dataset$rec_rate_total=(dataset$SCO1.sum+dataset$SCO2.sum)/(dataset$NCO1.sum+dataset$NCO2.sum+dataset$SCO1.sum+dataset$SCO2.sum)
```

##Recomb_figure_total

```{r}
#The base of the plot

Recomb_figure=ggplot(aes(y=rec_rate_total,x=Day..letter.of.vial., col=Treatment,label=total_offspring),data=dataset)+ylab("% recombination")+ggtitle("Total Recombination rate vs. Days post-mating")+theme_base()

#Now add points and change labels for x-axis

Recomb_figure=Recomb_figure+scale_colour_manual(values=c("blue", "red"))+geom_point(alpha=0.6,size=3)+scale_x_discrete(name="Day",labels=c("1-2","3-4","5-6","7-9","10-12","13-15"))

#Recomb_figure

#Add a line through the median of the points

Recomb_figure=Recomb_figure+stat_summary(fun.y = median, geom="line",aes(group=Treatment),size=2)

#Add a label for sample size

Recomb_figure=Recomb_figure+ggtitle("Total Recombination rate vs. Days post-mating") +geom_text(check_overlap = F,hjust = 0, nudge_x = 0.05,angle=45,size=3) +ylim(0,1.2)
Recomb_figure
```

##Statistical analysis
###the dataset necessary for the statistics is the cleaned up version of the summary data

```{r}
dataset2=read.csv("cleanedup_exp3.csv",header=T,stringsAsFactors = F)
dataset2$Treatment=as.character(dataset2$Treatment)
```

###get mean fecundity

```{r}
tapply(dataset2$fecundity,dataset2$Treatment,mean)
tapply(dataset2$fecundity,dataset2$Day,mean)
```

###poisson regression, similar to a t-test for count data

```{r}
fit=glm(fecundity~Treatment*Day..letter.of.vial.,data=dataset2,family=quasipoisson)
summary(fit)
```

###anova results for fecundity

```{r}
Experiment3_fecundity_stats=anova(fit, test="Chisq")
write.csv(Experiment3_fecundity_stats,"Experiment3_fecundity_anova.csv")
```

###post-hoc test for fecundity

```{r}
fit_contrast <- emmeans::emmeans(fit, "Treatment", by="Day..letter.of.vial.", mode="kenward-roger")
fit_contr <- contrast(fit_contrast, method="trt.vs.ctrl")
pheno_contr <- as.data.frame(summary(fit_contr))
pheno_contr
```
```{r}
write.csv(pheno_contr,"Experiment3_fecundity_posthoc.csv")
```

###Now do the same, but with recombination rate

###remove vials with few progeny

```{r}
dataset2=dataset2[dataset2$total_offspring>=10,]
```

###sum of crossovers in yv

```{r}
dataset2$num_CO_1=dataset2$SCO1.sum+dataset2$SCO2.sum
```

###sum of non-crossovers in yv

```{r}
dataset2$num_NCO_1=dataset2$total_offspring-(dataset2$SCO1.sum+dataset2$SCO2.sum)
```

###total recombination rate

```{r}
dataset2$rec_rate_yv=(dataset2$SCO1.sum+dataset2$SCO2.sum)/dataset2$total_offspring
```

###get mean recombination

```{r}
tapply(dataset2$rec_rate_yv,dataset2$Treatment,mean)
tapply(dataset2$rec_rate_yv,dataset2$Day,mean)
```

###to clean the dataset we get rid of the characters on the left and right of the F1 vial numbers

```{r}
dataset2$F1.Vial=as.numeric(gsub("[^0-9\\.]","",dataset2$F1.Vial))
 print(dataset2$F1.Vial)
```

###logistic regression, similar to a t-test for count data

```{r}
fit2=glmer(cbind(num_CO_1,num_NCO_1)~(1|F1.Vial)+Treatment*Day..letter.of.vial.,data=dataset2,
         family=binomial(link="logit"))
summary(fit2)
```

###anova for the recombination rates in vy region

```{r}
anova_vy=Anova(fit2,test="Chisq")
write.csv(anova_vy,"experiment3_recrate_anova.csv")
```
```{r}
fit_contrast_rec <- emmeans::emmeans(fit2, "Treatment", by="Day..letter.of.vial.", mode="kenward-roger")
fit_contr_rec <- contrast(fit_contrast_rec, method="trt.vs.ctrl")

pheno_contr_rec <- as.data.frame(summary(fit_contr_rec))
pheno_contr_rec
```


```{r}
write.csv(pheno_contr_rec,"Experiment3_recrate_vy_posthoc_table.csv")

#add in odds ratio and standard error
vy_or=exp(pheno_contr_rec$estimate)
vy_error=(pheno_contr_rec$SE)
x=cbind(vy_or,vy_error)
day=c("a","b","c","d","e","f")
x=cbind(day,x)
#convert p-values to stars for plot
vy_sig=ifelse(pheno_contr_rec$p.value<0.001,"***",ifelse(pheno_contr_rec$p.value<0.01,"**",ifelse(pheno_contr_rec$p.value<0.05,"*","")))
y=cbind(x,vy_sig)
odds_ratios=y
write.csv(y,"Experiment3_odds.csv")
```
```{r}
odds_ratios=read.csv("Experiment3_odds.csv", header=TRUE)

odds_figure=ggplot(aes(y=vy_or,x=day,group=1),data=odds_ratios)+scale_colour_manual(values=c("black"))+
  geom_point(alpha=0.6,size=3)+ylab("Odds Ratios")+theme_base()+geom_hline(yintercept = 1,linetype="dashed",color="grey")+ylim(0.5,2.6)+
  scale_x_discrete(name="Day",labels=c("1-2","3-4","5-6","7-9","10-12","13-15"))+geom_line()+geom_errorbar(aes(ymin=vy_or-vy_error,ymax=vy_or+vy_error))+
  annotate(geom="text", x=1, y=1.75, label=vy_sig[1],color="black",size=10)+annotate(geom="text", x=2, y=1.75, label=vy_sig[2],color="black",size=10)+annotate(geom="text", x=3, y=1.75, label=vy_sig[3],color="black",size=10)+annotate(geom="text", x=4, y=1.75, label=vy_sig[4],color="black",size=10)+annotate(geom="text", x=5, y=1.75, label=vy_sig[5],color="black",size=10)+annotate(geom="text", x=6, y=1.75, label=vy_sig[6],color="black",size=10)
odds_figure
```

##THE END







